MLPPI Wizard: An Automated Multi-level Partitioning Tool on Analytical Workloads

### 2017-03-10 mlppi-comsis17/paper.tex
New updates on ComSIS paper

In the subsequent step we gather the simplified predicates and 
then extract ranges from the collected predicates. 
This task can also be conducted through an API call to the server. 
The query predicate is of the form 
\begin{center}
\begin{tabular}{l} 
\form{\tt variable} \form{\tt op} \form{\tt constant},
\end{tabular}
\end{center}
where \form{\tt variable} is a field from {\tt LINEORDER}, 
\form{\tt op} is in 
\begin{center}
\begin{tabular}{l} 
\{{\tt $ = $ }, {\tt $ < $}, {\tt $ <= $}, {\tt $ >= $}, {\tt $ > $}, 
{\tt IN}\},
\end{tabular}
\end{center}
and \form{\tt constant} represents a constant value(s). 
All \form{\tt op}s are self-explanatory. 
In particular, `{\tt IN}' is a list predicate implying `{\tt OR}' operator.
In the running example, 
we can collect a set of predicates $P$ = \{$p$1,$p$2,$p$3,$p$4\} from $Q$, where

@@

In turn, the wizard constructs a bi-directional {\em map} between a query and 
associated predicates, so that for each query the corresponding predicate(s) can be 
found directly in the map, and vice versa. 
In the example, a map, called $M$1, is built 
between $Q$ and $P$, and $M$1 is filled with the following entries:

@@

Next, we gather fields referenced by predicates in $P$, extract 
ranges from the predicates, and group the ranges on each of the fields. 
Assume $I$ to be the list of fact table fields referenced by $P$. 
In the workload $Q$, we can see that there are two fields used by $P$: 
{\tt LO\_DISCOUNT}, {\tt LO\_QUANTITY}. 
These are added to $I$.

After that, we construct a set of distinct ranges on each field in $I$, using the predicates in $P$. 
Then, the whole range set becomes a kind of two-dimensional array, called $R$. 
The array $R$ captures the range representation of the predicates. 
Each entry under a field in $R$ represents a pair of (start and end) values forming the range. 
In the range representation, infinity ($\infty$) or -infinity (-$\infty$) can 
be used for unbounded ranges. 
``['' or ``]'' are used for closed ranges, and ``('' or ``)'' for open ranges. 
In particular, {\tt IN} predicate is represented by multiple ranges. 
For instance, the above $p$1 can be represented by the following three ranges 
\range{1, 1}, \range{4, 4}, \range{5, 5}; however, the last two consecutive 
ranges can be consolidated as \range{4, 5}. 
In the continuing example the range set $R$ is formed by the entries below.

@@

Then, each predicate can be mapped to its associated ranges in $R$. That is, 
we can have another map, called $M$2, between $P$ and $R$. 
Let $R$[$i$, $j$] denote the interval value for the $i$-th field and $j$-th range in $R$. 
$R$[1, 3], for instance, indicates range \ropenrange{7, \infty} under {\tt LO\_DISCOUNT}. 
In the running example, $M$2 is constructed as follows:

>>>>

In the subsequent step we gather the simplified predicates and 
then extract ranges from the collected predicates. 
This task can also be performed through an API call to the server. 
The query predicate is of the form 
\begin{center}
\begin{tabular}{l} 
\form{\tt variable} \form{\tt op} \form{\tt constant},
\end{tabular}
\end{center}
where \form{\tt variable} is a field (column) from {\tt LINEORDER}, 
\form{\tt op} is in 
\begin{center}
\begin{tabular}{l} 
\{{\tt $ = $ }, {\tt $ < $}, {\tt $ <= $}, {\tt $ >= $}, {\tt $ > $}, 
{\tt IN}\},
\end{tabular}
\end{center}
and \form{\tt constant} represents a constant value(s). 
All \form{\tt op}s are self-explanatory. 
In particular, `{\tt IN}' is a list predicate implying `{\tt OR}' operator.
In the running example, 
we can collect a set of predicates $P$ = \{$p$1,$p$2,$p$3,$p$4\} from $Q$, where

@@

Next, we construct a bi-directional {\em map} between a query and 
associated predicates, so that for each query the corresponding predicate(s) 
can be found directly in the map, and vice versa. 
In the example, we get a map, called $M$1, built between $Q$ and $P$. 
$M$1 gets filled with the following entries:

@@

For each field referenced by the predicates, 
we then build a set of distinct ranges on that field. 
Specifically, we gather fields referenced by predicates in $P$, extract 
ranges from the predicates, and group the ranges of each of the referenced fields. 
Assume $I$ to be the list of fact table fields referenced by $P$. 
In the workload $Q$ we add to $I$ two fields, {\tt LO\_DISCOUNT} and {\tt LO\_QUANTITY}, used by $P$. 
The whole range set then becomes a kind of two-dimensional array, called $R$, capturing 
the range representation of the gathered predicates. 
Each entry under a field in $R$ represents a pair of (start and end) values forming the range. 
In the range representation, infinity ($\infty$) or -infinity (-$\infty$) can 
be used for unbounded ranges. 
``['' or ``]'' are used for closed ranges, and ``('' or ``)'' for open ranges. 
In particular, {\tt IN} predicate is represented by multiple ranges. 
For instance, the above $p$1 can be represented by the following three ranges 
\range{1, 1}, \range{4, 4}, \range{5, 5}; however, the last two consecutive 
ranges can be consolidated as \range{4, 5}. 
In the continuing example the field-wise distinct range set ($R$) is formed by the entries below.

@@

We then build another map, called $M$2, between $P$ and $R$. 
That is, we associate each predicate in $P$ with a corresponding range in $R$. 
Let $R$[$i$, $j$] denote the interval value for the $i$-th field and $j$-th range in $R$. 
$R$[1, 3], for instance, indicates range \ropenrange{7, \infty} under {\tt LO\_DISCOUNT}. 
In the running example, $M$2 is constructed as follows:
